Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-04-16 thirdPaper/draft/typeConeAssociahedra.tex
Th 4.27 mutation conflations for Artin algebras

% PG et Yann
\newcommand{\field}{\mathbb{K}}
\newcommand{\cat}{\mathcal{C}}
\newcommand{\Hom}[1]{\operatorname{Hom}_{#1}}
\newcommand{\susp}{\Sigma}
\newcommand{\add}{\operatorname{add}}
\newcommand{\MOD}{\operatorname{mod}}
\newcommand{\End}[1]{\operatorname{End}_{#1}}
\newcommand{\spl}{\operatorname{sp}}
\newcommand{\Ksp}{K_0^{\spl}}
\newcommand{\Kg}{K_0^{\b{g}}}
\newcommand{\ind}{\operatorname{ind}}
\newcommand{\coker}{\operatorname{coker}}
\newcommand{\CC}{\phi}
\newcommand{\proj}{\operatorname{proj}}

@@

\begin{theorem}
\label{theo::brick-algebra-condition}
 Let~$\Lambda$ be a brick algebra which is finite-dimensional over a field~$\field$\pg{True for any Artin algebra?}.  Then the almost-split conflations of~$K^{[-1,0]}(\proj \Lambda)$ are mutation conflations if and only if for any indecomposable~$\Lambda$-module~$M$, the space~$\Hom{\Lambda}(M, \tau^2 M)$ vanishes.
\end{theorem}

@@

We have an exact sequence
\[
\Hom{K^b}(E,E) \to \Hom{K^b}(E,X) \to \Hom{K^b}(E,\susp \tau X) \to \Hom{K^b}(E,\susp E).
\]
The first map is surjective since~$E\to X$ is right almost split.
We can conclude by (4).
\end{proof}

\begin{lemma}
\label{lemm::if-AR-sequences-are-mutation}
 Let~$\Lambda$ be an Artin algebra.  Assume that for any almost-split conflation~$\tau X \infl E \defl X$ of~$K^{[-1,0]}(\proj\Lambda)$, we have that~$\Hom{K^b}(X, \susp E) = \Hom{K^b}(E, \susp X) = 0$ and~$\Hom{K^b}(\tau X, \susp E) = \Hom{K^b}(E, \susp \tau X) = 0$.  Then for any indecomposable~$\Lambda$-module~$M$, we have that~$\Hom{\Lambda}(M, \tau^2 M)$ vanishes.
\end{lemma}
\begin{proof}

\end{proof}

@@

We know describe a setup satisfying the assumptions of Section~\ref{setting::extricat} that, at the same time, subsumes all examples given so far, while being easily checked in practice.\yann{A reformuler...}

@@

One can check directly that the condition of Theorem~\ref{theo::brick-algebra-condition} is not satisfied. 
\end{example}

>>>>

% PG et Yann
\newcommand{\field}{\mathbb{K}}
\newcommand{\cat}{\mathcal{C}}
\newcommand{\Hom}[1]{\operatorname{Hom}_{#1}}
\newcommand{\susp}{\Sigma}
\newcommand{\add}{\operatorname{add}}
\newcommand{\MOD}{\operatorname{mod}}
\newcommand{\End}[1]{\operatorname{End}_{#1}}
\newcommand{\spl}{\operatorname{sp}}
\newcommand{\Ksp}{K_0^{\spl}}
\newcommand{\Kg}{K_0^{\b{g}}}
\newcommand{\ind}{\operatorname{ind}}
\newcommand{\coker}{\operatorname{coker}}
\newcommand{\CC}{\phi}
\newcommand{\proj}{\operatorname{proj}}
\newcommand{\bsm}{\begin{smallmatrix}}
\newcommand{\esm}{\end{smallmatrix}}

@@

\begin{theorem}
\label{theo::brick-algebra-condition}
 Let~$\Lambda$ be an Artin algebra all of whose non-injective indecomposable objects are bricks and all of whose indecomposable objects are~$\tau$-rigid\yann{Are there examples of representation-infinite algebras which satisfy those assumptions?}. Then the almost-split conflations of~$K^{[-1,0]}(\proj \Lambda)$ are mutation conflations if and only if for any non-projective indecomposable~$\Lambda$-module~$M$, the space~$\Hom{\Lambda}(M, \tau^2 M)$ vanishes.
\end{theorem}

\begin{remark}
Assume that~$\Lambda$ is a finite-dimensional brick algebra which is representation-finite.
Then~\cite[Theorem 6.2]{DemonetIyamaJasso} implies that the assumptions of Theorem~\ref{theo::brick-algebra-condition} are satisfied.
\end{remark}

@@

We have an exact sequence
\[
\Hom{K^b}(E,E) \to \Hom{K^b}(E,X) \to \Hom{K^b}(E,\susp \tau X) \to \Hom{K^b}(E,\susp E).
\]
The first map is surjective since~$E\to X$ is right almost-split.
We can conclude by (4).
\end{proof}

\begin{lemma}
Let~$\Lambda$ be an Artin algebra, and let~$Y\infl E\defl X$ be a conflation in~$K^{[-1,0]}(\proj\Lambda)$. Then:
\begin{enumerate}
 \item If~$\Hom{K^b}(Y,\susp E)=0$, we have~$\Hom{\Lambda}(FX,\tau FY)=0$.
 \item If moreover~$Z\notin\add(\susp\Lambda)$ and the conflation is almost-split, then~$\Hom{\Lambda}(FX,\tau^2 FX)=0$.
\end{enumerate}
\end{lemma}

\begin{proof}
By Lemma~\ref{lemm::AR-conflations-and-sequences}, \emph{(2)} follows from \emph{(1)}.
Let us prove \emph{(1)}.
The conflation induces an exact sequence~$\Hom{K^b}(Y,\susp E) \to \Hom{K^b}(Y,\susp X) \to \Hom{K^b}(Y,\susp^2,Y)$.
The complex~$Y$ being two-term, we have~$\Hom{K^b}(Y,\susp^2,Y)=0$.
Assuming that~$\Hom{K^b}(Y,\susp E)=0$, we deduce that~$\Hom{K^b}(Y,\susp X)=0$, and hence that~$\Hom{\Lambda}(FX,\tau FY) = 0$ by~\cite[Lemma 3.4]{AdachiIyamaReiten}.
\end{proof}

\begin{lemma}
Let~$\Lambda$ be an Artin algebra, and let~$P$ be an indecomposable projective~$\Lambda$-module which is a brick.
Let~$I$ be indecomposable injective with socle~$P/\operatorname{rad}P$, and~$\alpha$ be the composition~$P\to\operatorname{soc} I \to I$.
Then, composition with~$\alpha$ induces a surjection~$\Hom{\Lambda}(P,I)\to\Hom{\Lambda}(P,I)$.
\end{lemma}

\begin{proof}
{\bf Small gap}
\end{proof}

\begin{lemma}\label{lemm::initial meshes}
Let~$\Lambda$ be an Artin algebra, and let~$P$ be an indecomposable projective~$\Lambda$-module which is a brick.
Let~$I$ be indecomposable injective with socle~$P/\operatorname{rad}P$, and~$\alpha$ be the composition~$P\to\operatorname{soc} I \to I$.
Complete the morphism~$\alpha$ to a triangle~$P\to Y\to E\to \susp P$ in~$K^{[-1,0]}(\proj\Lambda)$, where~$Y$ is a minimal projective presentation of~$I$.
Then~$Y\infl E\defl \susp P$ is an Auslander--Reiten conflation in~$K^{[-1,0]}(\proj\Lambda)$, which is also a mutation conflation.
\end{lemma}

\begin{proof}
In this proof, we implicitely identify the projective indecomposable~$\Lambda$-module~$P$ with the complex~$P$ concentrated in degree~$0$.

We first show that the conflation is almost-split.

blahblahblah

We now check that the conflation~$Y\infl E\defl\susp P$ is a mutation conflation.

(0) Since~$P$ is concentrated in only one degree, it is rigid in~~$K^{[-1,0]}(\proj\Lambda)$ and so is~$\susp P$.
Moreover,~$FY=I$ is~$\tau$-rigid, hence~$Y$ is rigid in~$K^{[-1,0]}(\proj\Lambda)$ by~\cite[Lemma 3.4]{AdachiIyamaReiten}.

(1) We have~$\Hom{K^b}(E,\susp^2 P)=0$ since~$E$ is concentrated in degrees~$-1$ and~$0$.

(2) $\Hom{K^b}(Y,\susp E)=0$: There is an exact sequence
\[
\Hom{K^b}(Y,\susp Y) \to \Hom{K^b}(Y,\susp E) \to \Hom{K^b}(Y,\susp^2 P),
\]
where the first term vanishes by (0) and the last term vanishes since~$Y$ is in degrees~$-1, 0$ while~$\susp^2 P$ is concentrated in degree~$-2$.

(3) $\Hom{K^b}(E,\susp E)=0$: There is an exact sequence
\[
 \Hom{K^b}(\susp Y,\susp E) \to \Hom{K^b}(\susp P, \susp E) \to \Hom{K^b}(E,\susp E) \to \Hom{K^b}(Y,\susp E)
\]
in which the last term is zero by (2).
The first morphism is isomorphic to~$\Hom{K^b}(Y,E) \to \Hom{K^b}(P,E)$, which is surjective by the previous lemma.

(4) $\Hom{K^b}(E,\susp Y) = 0$: Consider the exact sequence
\[
 \Hom{K^b}(E,E) \to \Hom{K^b}(E,\susp P) \to \Hom{K^b}(E,\susp Y) \to \Hom{K^b}(E,\susp E).
\]
Its last term vanishes by (3) and its first morphism is surjective because~$E\to \susp P$ is right almost-split.

(5) $\Hom{K^b}(\susp P,\susp E= 0$:
Let~$P\xrightarrow{f} E$ be any morphism.
Since~$P$ is rigid,~$f$ factors through the inflation~$\beta$ in the conflation~$Y\infl E\defl \susp P$, and there is some morphism~$g$ such that~$f=\beta g$.
By the previous lemma, or by (4), the morphism~$g$ factors through~$\alpha$: there is some~$h$ such that~$g= h \alpha$.

{\bf Small gap:} There is an~$h'$ such that~$h \alpha = \alpha h'$ and we deduce that~$f=\beta h \alpha = \beta \alpha h' = 0$.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{theo::brick-algebra-condition}]
 The theorem follows from the slightly more general statements of Lemma~\ref{lemm::then-AR-seq-are-mutation} and Lemma~\ref{lemm::initial meshes}.
\end{proof}

% \begin{lemma}
% \label{lemm::if-AR-sequences-are-mutation}
%  Let~$\Lambda$ be an Artin algebra.  Assume that for any almost-split conflation~$\tau X \infl E \defl X$ of~$K^{[-1,0]}(\proj\Lambda)$, we have that~$\Hom{K^b}(X, \susp E) = \Hom{K^b}(E, \susp X) = 0$ and~$\Hom{K^b}(\tau X, \susp E) = \Hom{K^b}(E, \susp \tau X) = 0$.  Then for any indecomposable~$\Lambda$-module~$M$, we have that~$\Hom{\Lambda}(M, \tau^2 M)$ vanishes.
% \end{lemma}
% \begin{proof}
%  
% \end{proof}

@@

We now describe a setup satisfying the assumptions of Section~\ref{setting::extricat} that, at the same time, subsumes all examples given so far, while being easily checked in practice.\yann{A reformuler...}

@@

One can check directly that the condition of Theorem~\ref{theo::brick-algebra-condition} is not satisfied.
For example, the module~$\bsm2\\1\esm$ is~$\tau$-rigid since~$\tau(\bsm2\\1\esm) = \bsm3\\1\esm$, while~$\Hom{\Lambda}\left(\bsm2\\1\esm,\tau^2(\bsm2\\1\esm)\right) = \Hom{\Lambda}(\bsm2\\1\esm,\bsm1\\2\esm) \neq 0$.
\end{example}
